---
title: "Assignment 1 : Linear Regression"
output: html_document
date: "2023-11-29"
---


Importing packages
```{r, message = FALSE, warning=FALSE}
library(dplyr)
library(readr)
library(knitr)
library(kableExtra)
library(DT)
library(ggplot2)
library(ggmice)
```
Reading in the datasets
```{r, message = FALSE}
n_em <- read_csv("data/nitrous-oxide-emissions.csv", show_col_types = FALSE) %>%
  filter(Year == 2020)

#n_em_per_cap <- read_csv("data/per-capita-nitrous-oxide.csv") %>%
  #filter(Year == 2020)
#n_em_per_cap

n_em_per_sec <- read_csv("data/per-capita-nitrous-oxide-sector.csv") %>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
pop_den <- read_csv("data/population-density.csv")%>%
  filter(Year == 2020) %>% 
  select(-Year,- Code)
forest_area <- read_csv("data/population-density.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
gdp_per_cap <- read_csv("data/gdp-per-capita-worldbank.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
#meat_supp_per_cap <- read_csv("data/meat-supply-per-person.csv")%>%
 # filter(Year == 2020)%>% 
  #select(-Year,- Code)
n_fert_app_per_hect <- read_csv("data/nit_fert_app.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
agr_land <- read_csv ("data/agricultural-land.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
ann_export <- read_csv("data/exports-of-goods-and-services-constant-2010-us.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
ann_import <- read_csv("data/imports-of-goods-and-services-constant-2010-us.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
urban_v_rural <- read_csv("data/urban-vs-rural-majority.csv")%>%
  filter(Year == 2020)%>% 
  select(-Year,- Code)
water_resc <- read_csv("data/renewable-water-resources-per-capita.csv")%>%
  filter(Year == 2019)%>% 
  select(-Year,- Code)
wetlands <- read_csv("data/coverage-of-wetlands.csv")%>%
   filter(Year == 2017)%>% 
  select(-Year,- Code)
```

Merging the datasets
```{r, message = FALSE}
df <- n_em %>%
  left_join(pop_den, by = "Entity") %>%
  left_join(n_em_per_sec, by = "Entity")%>%
  left_join(forest_area, by = "Entity")%>%
  left_join(gdp_per_cap, by = "Entity") %>%
  left_join(n_fert_app_per_hect, by = "Entity") %>%
  left_join(agr_land, by = "Entity")%>%
  left_join(ann_export, by = "Entity") %>%
  left_join(ann_import, by = "Entity")%>%
  left_join(urban_v_rural, by = "Entity")%>%
  left_join(water_resc, by = "Entity")%>%
  left_join(wetlands, by = "Entity")%>%
  select(-"Population density.x", - "Year")

```

Entities Filter 
```{r, message = FALSE, results='hide'}

df %>% filter(!Entity %in% c(
    "Africa", "America", "North America", "South America", "Asia", "Europe", "Oceania",
    "High-income countries","Upper-middle-income countries", "Low-income countries", "Lower-middle-income countries",
    "European Union (27)", "Asia (excl. China and India)", 
    "Europe (excl. EU-27)", "Europe (excl. EU-28)", 
    "European Union (28)", "Least developed countries (Jones et al. 2023)", 
    "North America (excl. USA)", "OECD (Jones et al. 2023)", "World"))
df

```
Renaming some columns 
```{r}
colnames(df)[1:17] <- c("Country", 
                        "Code",
                        "Annual N20 Emmissions ( Co2 Equiv.)" , 
                        "Agriculture Waste Emissions", 
                        "Waste Emissions", 
                        "Land Use Change and Forestry Emissions", 
                        "Energy Production Emissions", 
                        "Population Density", 
                        "GDP per Capita", 
                        "Nutrient nitrogen use per area of cropland",
                        "Agricultural Land Use",
                        "Exports of Goods and Services", 
                        "Imports of Goods and Services", 
                        "Urban Population %",
                        "Rural Population %",
                        "Freshwater Resources per Capita", 
                        "Wetlands area")


```
Creating a categorical column based on the predominant nitrogen emission sector 
```{r}
df %>% 
  mutate("Highest Emission Sector" = case_when(
    "Agriculture Waste Emissions" >
      "Waste Emissions" & "Agriculture Waste Emissions" > 
      "Land Use Change and Forestry Emissions" & "Agriculture Waste Emissions" > 
      "Energy Production Emissions" ~ 
      "Agriculture Waste",
    "Waste Emissions" > 
      "Agriculture Waste Emissions" & "Waste Emissions" > 
      "Land Use Change and Forestry" & "Waste Emissions" > 
      "Energy Production Emissions" ~ 
      "Waste Emissions",
    "Land Use Change and Forestry Emissions" > 
      "Agriculture Waste Emissions" & "Land Use Change and Forestry Emissions" > 
      "Waste Emissions" & "Land Use Change and Forestry Emissions" > 
      `Energy Production Emissions` ~ 
      "Land Use Change and Forestry",
    `Energy Production Emissions` > 
      "Agriculture Waste Emissions" & "Energy Production Emissions" > 
      "Land Use Change and Forestry Emissions" & "Energy Production Emissions" > 
      "Waste Emissions" ~ 
      "Energy Production",
    TRUE ~ NA_character_
  ))
```

Creating a binary variable depending on whether the majority of population lives in urban or rural areas
```{r}
df %>%
  mutate("Majority Population( Urban/Agricultural" = case_when(
    "Urban Population %" > "Rural Population %" ~ "Urban",
    "Rural Population %" > "Urban Population %" ~ "Rural"
  ))
```
Removing rows with missing values 
```{r}
#df <- na.omit(df)
#df

```

```{r}
df
```

Applying normalization to some variables to facilitate easier interpretation through plotting 

```{r}
#Annual N20 Emmissions
df_normalized <- df
df_normalized$"Annual N20 Emmissions ( Co2 Equiv.)" <- as.numeric(as.character(df_normalized$"Annual N20 Emmissions ( Co2 Equiv.)"))
df_normalized$"Annual N20 Emmissions ( Co2 Equiv.)" <- log(df_normalized$"Annual N20 Emmissions ( Co2 Equiv.)")

#Land Use and Forestry (min_max)

#min_max_normalize <- function(x){(x - min(x)/max(x) - min(x))}
#df_normalized$`Land Use Change and Forestry Emissions` <- min_max_normalize(df_normalized$`Land Use Change and Forestry Emissions`)
#df_normalized$`Land Use Change and Forestry Emissions` <- as.numeric(df_normalized$`Land Use Change and Forestry Emissions`)
#df_normalized$`Land Use Change and Forestry Emissions ` <- log(df_normalized$`Land Use Change and Forestry Emissions`)

# Agricultural Land Use 
df_normalized$`Agricultural Land Use` <- log(df_normalized$`Agricultural Land Use`)
df_normalized$`Exports of Goods and Services` <- log(df_normalized$`Exports of Goods and Services`)
df_normalized$`Imports of Goods and Services` <- log(df_normalized$`Imports of Goods and Services`)
df_normalized$`Freshwater Resources per Capita` <- log(df_normalized$`Freshwater Resources per Capita`)
```


# Now, we plot all variables with nitrous oxide emissions in an attempt to detect possible trends

# Nitrous oxide emissions vs Renewable freshwater resources

```{r}
N2OvsFreshwater <- ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
   y = `Freshwater Resources per Capita`
)) +
geom_point() + 
ylim(0, max(df_normalized$`Freshwater Resources per Capita`))
print(N2OvsFreshwater)

```

# Nitrous oxide emissions vs Wetlands area

```{r}

N2OvsWetlands <- ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Wetlands area`)) +
  geom_point() +
  ylim(0, 65)

print(N2OvsWetlands)

```

# We have to filter out outliers in order to obtain readable data.

Firstly, we must calculate the IQR for wetlands area:

```{r}
Q1 <- quantile(df_normalized$`Wetlands area`, 0.25, na.rm = TRUE)
Q3 <- quantile(df_normalized$`Wetlands area`, 0.75, na.rm = TRUE)
IQR <- Q3 - Q1
```

Then, we define the lower and upper bounds

```{r}
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
```

Then, we filter out the outliers

```{r}
df_filtered <- df_normalized %>%
  filter(`Wetlands area` >= lower_bound & 
         `Wetlands area` <= upper_bound)
```

Finally, we can obtain readable plots:

```{r}
N2OvsWetlands <- ggplot(df_filtered, aes(x = `Annual N20 Emmissions ( Co2 Equiv.)`, y = `Wetlands area`)) +
  geom_point()
print(N2OvsWetlands)
```

From now on, removing outliers will be a standard procedure for each plot.

# Nitrous Oxide emissions vs Rural population %

```{r}
Q1_rural <- quantile(df_normalized$`Rural Population %`, 0.25, na.rm = TRUE)
Q3_rural <- quantile(df_normalized$`Rural Population %`, 0.75, na.rm = TRUE)
IQR_rural <- Q3_rural - Q1_rural

lower_bound_rural <- Q1_rural - 1.5 * IQR_rural
upper_bound_rural <- Q3_rural + 1.5 * IQR_rural

df_filtered <- df_normalized %>%
  filter(`Rural Population %` >= lower_bound_rural & 
         `Rural Population %` <= upper_bound_rural)

N2OvsRural <- ggplot(df_filtered, aes(x = `Annual N20 Emmissions ( Co2 Equiv.)`, y = `Rural Population %`)) +
  geom_point() 

print(N2OvsRural)
```

# N2O vs Urban population %

```{r}

Q1_urban <- quantile(df_normalized$`Urban Population %`, 0.25, na.rm = TRUE)
Q3_urban <- quantile(df_normalized$`Urban Population %`, 0.75, na.rm = TRUE)
IQR_urban <- Q3_urban - Q1_urban

lower_bound_urban <- Q1_urban - 1.5 * IQR_urban
upper_bound_urban <- Q3_urban + 1.5 * IQR_urban

df_filtered <- df_normalized %>%
  filter(`Urban Population %` >= lower_bound_urban & 
         `Urban Population %` <= upper_bound_urban)

N2OvsUrban <- ggplot(df_filtered, aes(x = `Annual N20 Emmissions ( Co2 Equiv.)`, y = `Urban Population %`)) +
  geom_point()

print(N2OvsUrban)
```

```{r}

df_normalized
```
Lets look at the relationships between N20 Emmissions and other variables 
```{r}
names(df_normalized)
# Agricultural Waste Emmissions 

N2vsAgriWaste <-
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Agriculture Waste Emissions`)) +
  geom_point() +
  ylim(0, 5
       )
  
N2vsAgriWaste 

#Waste Emmissions

N2vsWasteEm <-
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y =  `Waste Emissions`)) +
  geom_point() +
  ylim(0, 1
       )
N2vsWasteEm

#Land Use Change and Forestry Emissions 
  
N2vsUseForest <-
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Land Use Change and Forestry Emissions`)) +
  geom_point() +
  ylim(0, 0.05
       )
N2vsUseForest

#Energy Production Emmissions 

N2vsEnergy <-
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Energy Production Emissions`)) +
  geom_point() +
  ylim(0, 0.025
       )
N2vsEnergy

#Population Density 

N20vsPopDen <- 
   ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Population Density`)) +
  geom_point() +
  ylim(0, 750
       )

N20vsPopDen

#GDP per Capita 

N20vsGDP <- 
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `GDP per Capita`)) +
  geom_point() +
  ylim(0, 750
       )
#Nutrient Nitrogen Use 

N20vsNutrient <- 
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Nutrient nitrogen use per area of cropland`)) +
  geom_point() +
  ylim(0, max(df_normalized$`Nutrient nitrogen use per area of cropland`)
  )

N20vsNutrient

#Agricultural Land Use

N20vsAgriLand <- 
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Agricultural Land Use`)) +
  geom_point() +
  ylim(0, max(df_normalized$`Agricultural Land Use`)
  )

N20vsAgriLand

#Export
N20vsExports <- 
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Exports of Goods and Services`)) +
  geom_point() +
  ylim(0, max(df_normalized$`Exports of Goods and Services`)
  )

N20vsExports

#Imports
N20vsImports <- 
  ggplot(data = df_normalized, aes(
  x = `Annual N20 Emmissions ( Co2 Equiv.)`,
  y = `Imports of Goods and Services`)) +
  geom_point() +
  ylim(0, max(df_normalized$`Imports of Goods and Services`)
  ) +
  xlim(0, max(df_normalized$`Annual N20 Emmissions ( Co2 Equiv.)`))

N20vsImports
```
